---
title: "R Data Management"
author: "Steven Nanga"
date: "2024-02-29"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Data Cleaning and Preprocessing:
- **Handling Missing Values:**
  - Identify missing values (NA, NaN, etc.) and decide how to handle them.
  - Use functions like na.omit(), complete.cases(), or impute missing values.
  
```{r}
# cleaned_data = na.omit(sales_data_sample)

```
  
- **Removing Duplicates:**
  - Detect and remove duplicate rows using duplicated() and unique().  
  - Identify duplicates using duplicated():
  
```{r}
#duplicated_rows <- duplicated(sales_data_sample)
# print(cleaned_data, n= 100 )
# anyDuplicated(sales_data_sample)
# anyDuplicated(cleaned_data)
 

```
  
  - Remove duplicates using unique():
  
```{r}
# unique_data = unique(cleaned_data)
# print(unique_data)

```
 
- **Data Transformation:**
  - Convert data types (e.g., numeric to character, date formatting).
  - Create new variables using existing ones (e.g., calculating percentages). 
  
- **Outlier Detection and Treatment:**
  - Identify outlines using box plots, histograms, or statistical methods.
  
```{r}
# boxplot(cleaned_data$QUANTITYORDERED, main = "Quantity Ordered")
# hist(cleaned_data$PRICEEACH, main = "Quantity Ordered")

```
  - Decide whether to remove, transform, or keep outlines. 

### Data Aggregation and Summarization:
- **Grouping Data:**
  - Use group_by() to group data by one or more variables.
  - Aggregate data using functions like summarize(), count(), or mean().
  
```{r}
#library(dplyr)
#summarized_data <- cleaned_data %>%
#  group_by(PRODUCTLINE) %>%
#   summarize(AvgPrice = mean(PRICEEACH))
 
# print(summarized_data)

```
  
- **Pivoting Data:**
  - Convert data from long to wide format (and vice versa) using pivot_longer() and pivot_wider(). 
  
```{r}
# wide_data <- original_data %>%
 # pivot_wider(names_from = Category, values_from = Value)

```
  
### Data Joins and Merging:
- **Merging Datasets:**
  - Combine data from different sources using functions like merge(), join(), or   bind_rows().
  - Understand different join types (inner, left, right, full).
  
```{r}
#merged_data <- merge(Sales, Customers, by = "CustomerID", all = TRUE)

```
- **Reshaping Data:**
  - Transform data between wide and long formats using melt() and cast().
  
```{r}
# install.packages("reshape2")
# library(reshape2)
# long_data <- melt(wide_data, id.vars = "Date")

```
  
### Data Validation and Quality Assurance:
- **Checking Data Integrity:**
  - Validate data against business rules or constraints.
  - Ensure consistency and accuracy.
  
```{r}
# valid_dates <- original_data$Date >= "2022-01-01" & original_data$Date <= "2022-12-31"

```
  
  