---
title: "R Fundamentals"
author: "Steven Nanga"
date: "2024-02-27"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Basic Syntax

### Assignment Operator (<- or =):
- Use the assignment operator to assign values to variables. For example


```{r}
x <- 10
y <- "Hello, R!"
z = 11

```

### Variables:
- Variables store data. You can name them as you like (following rules like starting with a letter or a dot).

```{r}
my_age <- 30
my_name <- "Alice"

```

### Comments: 
- Add comments to explain your code. Comments start with #. They are ignored by R but help others (and your future self) understand your code:

```{r}
# This is a comment
```

### Arithmetic Operators:
- R supports standard arithmetic operations:

```{r}
addition_result <- 5 + 3
subtraction_result <- 10 - 2
multiplication_result <- 4 * 6
division_result <- 20 / 5

```

### Logical Operators:
- These operators allow you to perform logical comparisons and combine conditions. Here are the key ones:
  1. **Logical AND (& or &&):**
     - The & operator checks if both conditions are true.
     - The && operator evaluates only the first condition and proceeds to the second condition         only if the first is true
     - Example:
```{r}
x <- 5
y <- 10
result1 <- x > 3 & y < 15  # TRUE
result2 <- x > 7 && y < 15  # FALSE

```

  2. **Logical OR (| or ||):**
     - The | operator checks if at least one condition is true.
     - The || operator evaluates only the first condition and proceeds to the second condition        only if the first is false.
     - Example:
     
```{r}
a <- TRUE
b <- FALSE
result3 <- a | b  # TRUE
result4 <- a || b  # TRUE
```
    
### Comparison Operators
- These operators compare values and return logical results.
- Common comparison operators:
   - \>: Greater than
   - \<: Less than
   - \>=: Greater than or equal to
   - \<=: Less than or equal to
   - \==: Equal to
   - \!=: Not equal to
- Example:

```{r}
num1 <- 7
num2 <- 10
result6 <- num1 > num2  # FALSE
result7 <- num1 == num2  # FALSE

```

## Data Types

### R Data Types

1. **Numeric (Double):**
   - Represents real numbers, including integers and decimals.
```{r}
numeric_value <- 3.14
```
   
2. **Integer:**
   - Represents whole numbers (no decimal part).
   - You can use the L suffix to denote an integer explicitly 
   
```{r}
integer_value <- 42
```
  
3. **Logical (Boolean):**
   - Takes values of TRUE or FALSE.
   - Often used for comparisons and conditional statements.
   
```{r}
logical_value <- TRUE
```

4. **Complex:**
   - Represents complex numbers with both real and imaginary parts.
   - Written as a + bi, where a and b are real numbers.
  
```{r}
complex_value <- 1 + 2i
```
  
5. **Character (String):**
   -Represents text data (words, sentences, etc.).
   -Enclosed in double or single quotes.
   
```{r}
character_value <- "Hello, world"
```
   
### Checking Data Types:
- Use the class() function to determine the data type of a variable:   

```{r}
x <- 10.5
class(x)  # Returns "numeric"

y <- "Hello, R!"
class(y)  # Returns "character"

```

###  Data type conversion in R

Data type conversion allows you to transform one type of data into another. Here are some common scenarios and functions for type conversion:

  1. **Numeric to Integer:**
      - Use as.integer() to convert a numeric value to an integer:
      
```{r}
num <- 42.5
int_num <- as.integer(num)  # Converts 42.5 to 42

```
      
  2. **Character to Numeric:**
      - Use as.numeric() to convert a character (string) to a numeric value:
      
```{r}
str_num <- "123.45"
numeric_str <- as.numeric(str_num)  # Converts "123.45" to 123.45

```

  3. **Character to Logical:**
      - Use as.logical() to convert a character to a logical value (TRUE or FALSE):
      
```{r}
str_logical <- "TRUE"
logical_value <- as.logical(str_logical)  # Converts "TRUE" to TRUE

```
   
  4. **Logical to Numeric:**
      - Logical values can be converted to numeric (0 for FALSE, 1 for TRUE): 
      
```{r}
my_logical <- TRUE
numeric_logical <- as.numeric(my_logical)  # Converts TRUE to 1

```
      
  5. **Factor to Character:**
      - Factors are categorical variables. Convert them to character for text representation:
      
```{r}
my_factor <- factor("High")
char_factor <- as.character(my_factor)  # Converts factor to "High"

```
      
  6. **Date and Time Conversions:**
     - Use functions like as.Date(), as.POSIXct(), or as.POSIXlt() to convert date and time formats.
     
## Data Structures

