---
title: " Installing and Loading R Packages"
author: "Steven Nanga"
date: "2024-02-28"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Installing R Packages:

- **What is an R Package?**
  - An R package is a collection of functions, data sets, documentation, and tests. It        allows you to organize your work and share it with others.
  - Packages enhance R by improving existing functionalities or adding new ones.
  - The DESCRIPTION file provides basic information about a package, including its       purpose, author, version, license, and dependencies.
  
- **Where to Find Packages?**
  - Packages are available from various sources:
     - **CRAN (Comprehensive R Archive Network):** The official repository with thousands of packages.
     - **CRAN Mirrors:** Alternative servers hosting CRAN packages.
     - **Bioconductor:** Specialized for biology-related packages.
     - **GitHub:** For developing versions of packages.
     
- **How to Install Packages?**
  - Use the install.packages() function:
```{r}
# install.packages("package_name")

```
  
  - For example, to install the dplyr package:

```{r}
# install.packages("dplyr")

```

### Loading R Packages:
- Using the library() Function:
  - Load a package into your R session:
```{r}
# library(package_name)

```
  - For example, to load the dplyr package:
  
```{r}
library(dplyr)
```
  
- Alternative: require() Function:
  - Similar to library(), but returns a logical value (useful for conditional checks).
  

- Checking Package Version:
  - Sometimes specific package versions are required for compatibility.
  - Ensure you have the correct version using:
  
```{r}
# packageVersion("package_name")

```
  
- Detaching a Package:
  - To unload a package from the current session:  
```{r}
# detach(package:package_name, unload = TRUE)

```
 
## Further Learning Resources

1. **R Tutorial:** http://openintrostat.github.io/oilabs-tidy/01_intro_to_r/intro_to_r.html
2. **Hands-on Programming with R:** https://github.com/rstudio-education/hopr
3. **All About R:** https://www.datacamp.com/blog/category/r-programming
4. **R Tutorial by GeeksforGeeks:** https://www.geeksforgeeks.org/r-tutorial/
5. **R for Data Science by Bookdown:** https://bookdown.org/roy_schumacher/r4ds/
6. **Statistics Globe:** https://statisticsglobe.com/
7. **Dataquest:** https://www.dataquest.io/  
  